---
title: "Health Services Research Journal Club"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---
### 11/05/2021: Health Recommendations and Selection

#### Paper of the day (Oster 2020)
Health Recommendations and Selection in Health Behaviors. AER: Insights (2020) https://doi.org/10.1257/aeri.20190355

#### Example of the day:
Pinnaple consumption and cardiovascular fitness. (simulated data)

```{r message=FALSE}
# Load the needed libraries
library(tidyverse)
library(ggpubr)
library(tableone)


# Load the data
data <- read.csv("https://raw.githubusercontent.com/Miquel-Serra/HSRdata/main/dataset2.csv")

# Inspect data
str(data)

# Descriptive Table
CreateTableOne(vars = c("weekly_sport_score","weekly_diet_score","weekly_sleep_score","pinnaple_consumption"),strata = "year",data)

# model 0
fit0 <- lm(cardiovascular_fitness ~ pinnaple_consumption, data)
summary(fit0)

# model 1
fit1 <- lm(cardiovascular_fitness ~ pinnaple_consumption:factor(year), data)
summary(fit1)

# model 2
fit2 <- lm(cardiovascular_fitness ~ pinnaple_consumption:factor(year) + weekly_sport_score, data)
summary(fit2)

# model 3
fit3 <- lm(cardiovascular_fitness ~ pinnaple_consumption:factor(year) + weekly_sport_score + weekly_diet_score, data)
summary(fit3)

# model 4
fit4 <- lm(cardiovascular_fitness ~ pinnaple_consumption:factor(year) + weekly_sport_score + weekly_diet_score + weekly_sleep_score, data)
summary(fit4)

# Coefficient plots
f1 <- data.frame(summary(fit1)$coefficients[2:4,])
f2 <- data.frame(summary(fit2)$coefficients[3:5,])
f3 <- data.frame(summary(fit3)$coefficients[4:6,])
f4 <- data.frame(summary(fit4)$coefficients[5:7,])
f1$model <- "basic"
f2$model <- "Control 1"
f3$model <- "Control 2"
f4$model <- "Control 3"
fits <- rbind(f1,f2,f3,f4)
fits$x <- rep(c(2006,2007,2008),4)
fits %>% ggplot(aes(x=x,y=Estimate,color=model)) + 
  geom_point() +
  theme_pubclean() +
  scale_color_viridis_d() +
  geom_errorbar(aes(ymin=Estimate-1.96*Std..Error,ymax=Estimate+1.96*Std..Error),width=.1)

## Bias
summary(lm(weekly_sport_score ~ pinnaple_consumption, data))
summary(lm(weekly_diet_score ~ pinnaple_consumption, data))
summary(lm(weekly_sleep_score ~ pinnaple_consumption, data))
summary(lm(cardiovascular_fitness ~ weekly_sport_score, data))
summary(lm(cardiovascular_fitness ~ weekly_diet_score, data))
summary(lm(cardiovascular_fitness ~ weekly_sleep_score, data))


```


### 20/04/2021: Risk-adjustment with claims data

#### Paper of the day (Wennberg 2013)
Observational intensity bias associated with illness adjustment: cross sectional analysis of insurance claims
BMJ 2013; 346 doi: https://doi.org/10.1136/bmj.f549 (Published 21 February 2013)

#### Example of the day:
Oncology: NSLCC stage 4 patients mortality. (simulated data)

a) what are the crude differences in mortality accross regions?
$$y_{i,j} = \beta_j x_{i,j} + \epsilon_{i,j}$$

where $y$ is the outcome of interest of patient $i$ and region $j$ and $\beta$ is the region-specific coefficient, $x$ is a matrix of dummies and $\epsilon$ is the residual.

b) what are the crude differences in mortality accross regions after controlling for age and sex of the patients?
$$y_{i,j} = \beta_j x_{i,j} + (age)_i + (sex)_i + \epsilon_{i,j}$$

c) what are the crude differences in mortality accross regions after controlling for age, sex, complexity and intensity of visits of the patients?
$$y_{i,j} = \beta_j x_{i,j} + (age)_i + (sex)_i + (complexity)_i + (intensity)_i + \epsilon_{i,j}$$


```{r message=FALSE}
# Load the needed libraries
library(tidyverse)
library(ggpubr)

# Load the data
data <- read.csv("https://raw.githubusercontent.com/Miquel-Serra/HSRdata/main/data_session_1.csv")

# Inspect data
str(data)

# Histogram
data %>% 
  group_by(region) %>% 
  summarise(mortality = mean(death)) %>%
  ggplot(aes(x=mortality*100)) + 
  geom_histogram(bins=50,fill="white",color="gray7")+
  theme_pubclean() + 
  xlab("Crude mortality rate (%)")

# Reduce units with few observations & palliative centers
exclud <- data %>% group_by(region) %>% summarise(n=n(),m=mean(death,na.rm=T)) %>% subset(n>30&m<.6&m>0)
data <- subset(data, region %in% exclud$region)

# Raw unadjusted
fit_un <- lm(death ~ factor(region),data)
f1<-data.frame(summary(fit_un)$coefficients)[2:dim(data.frame(summary(fit_un)$coefficients))[1],]
f1$region <- rownames(f1)
f1$region <- extract_numeric(f1$region)
f1<-f1[order(f1$Estimate),]
f1$x <- 1:dim(f1)[1]

f1 %>% ggplot(aes(x=x,y=Estimate*100)) +
  geom_point()+
  geom_errorbar(width=.01,aes(ymin=(Estimate-1.96*Std..Error)*100,ymax=(Estimate+1.96*Std..Error)*100)) +
  theme_pubclean()+
  ylab("Estimate mortality (%)") +
  coord_cartesian(ylim = c(-30,30))

# Adjusting for age and sex
fit_a1 <- lm(death ~ factor(region)+(age)+factor(sex),data)
f2<-data.frame(summary(fit_a1)$coefficients)[2:(dim(f1)[1]+1),]
f2$region <- rownames(f2)
f2$region <- extract_numeric(f2$region)
f2<-f2[ order(match(f2$region, f1$region)), ]
f2$x <- 1:dim(f2)[1]

# Adjusting for age and sex + additionals
fit_a2 <- lm(death ~ factor(region)+(age)+factor(sex)+scale(complex)+factor(intensity),data)
f3<-data.frame(summary(fit_a2)$coefficients)[2:(dim(f1)[1]+1),]
f3$region <- rownames(f3)
f3$region <- extract_numeric(f3$region)
f3<-f3[ order(match(f3$region, f1$region)), ]
f3$x <- 1:dim(f3)[1]

# Combining models
f1$model <- "a) unadjusted"
f2$model <- "b) age-sex adjusted"
f3$model <- "c) fully adjusted"

fits <- rbind(f1,f2,f3)
fits$region <- extract_numeric(fits$region)

pd <- position_dodge(0.5)
fits %>% ggplot(aes(x=x,y=Estimate*100,color=model)) +
  geom_point(position = pd)+
  ylab("Estimate mortality (%)")+
  theme_pubclean() +
  scale_color_viridis_d() +
  coord_cartesian(ylim = c(-30,30))

fits %>% ggplot(aes(x=x,y=Estimate*100,color=model)) +
  geom_errorbar(width=.001,position = pd,aes(ymin=(Estimate-1.96*Std..Error)*100,ymax=(Estimate+1.96*Std..Error)*100)) +
  geom_point(position = pd)+
  ylab("Estimate mortality (%)")+
  theme_pubclean() +
  scale_color_viridis_d() +
  coord_cartesian(ylim = c(-30,30))

rownames(fits) <- NULL
subset(fits,x%in%c(1))
subset(fits,x%in%c(53))


# R2
summary(fit_un)$r.squared
summary(fit_a1)$r.squared
summary(fit_a2)$r.squared

#  Fully Adjusted re-rank

fits %>% ggplot(aes(x=Estimate, fill=model))+
  geom_histogram(alpha=.3,bins = 40) + 
  theme_pubclean() +
  scale_fill_viridis_d() +
  geom_vline(xintercept = 0,linetype="dashed",alpha=.4)

```
